A Survey of Current Exoskeletons\\ and Their Control Architectures and Algorithms\\ (Draft 3.2)

### 2015-09-30 exos/exosText_MIT.tex
updated ass. and rec. on travers sections

 The MIT hardware and control system present a very well designed and thought-out system that focuses primarily on efficiency in design.  The series elastic linear ball screw hardware design seems to be a very smart way to implement efficient drives as well as to perform high fidelity force control.  The use of a strain gauge in the shank was also a novel means of measuring ground reaction forces.  This design may offer benefits in terms of repeatability over different operators as foot fall patterns for different users may include significant variations.

 The details on the controllers were somewhat sparse in the sense that, at least in \cite{}, the exact means through which the added torques were derived was not provided.  Overall though, the controllers were intelligently designed in the sense that they explicitly took the biomechanics of walking into account in their design and implementation.  A similar approach for a variety of different behaviors may be very beneficial to the TALOS project.

>>>>

 The MIT hardware and control system present a well designed system that focuses primarily on efficiency in design.  We generally support the series-elastic linear ball screw hardware design for their mechanical efficiency as well as ability to perform high fidelity force control.  Although, these drives will again be limited by bandwidth issues in highly-dynamic maneuvers.  The use of a strain gauge in the shank was also a novel means of measuring ground reaction forces.  This design may offer benefits in terms of repeatability over different operators, as foot fall patterns may include significant variations for different users or even in the same user in different environments.

 The details on the controllers were somewhat sparse in the sense that, at least in \cite{mit_control_2006}, the exact means through which the added torques discussed above were derived was not provided.  Overall, we support the design of the controllers in the sense that they explicitly took the biomechanics of walking into account in their design and implementation.  A similar approach for a variety of different behaviors may be very beneficial.

### 2015-09-30 exos/exosText_bodyExt.tex
updated ass. and rec. on travers sections

The Body Extender suit is one of the only full-body exoskeletons available today, and as such is a somewhat remarkable system. In terms of how the Body Extender system is designed and how it is controlled, the suit may be limited. Clearly, for the task of handling heavy objects a fully actuated suit may be the best option, but, generally speaking, this seems almost like an over engineered solution.  The Body Extender is a heavy, overly clunky suit.  This suit may be alright for quasi-static operations, but it is unlikely that it will be useful for more than pick and place type operations.  Again, seeing as this is basically what it was designed for, that seems alright.

In terms of the controllers highlighted in \cite{body_control_2012}, there are some good ideas.  For example, using a regressor to estimate unknown external forces may be an idea which is directly applicable to the TALOS project.  In the case of the Body Extender this regression is dependent on knowing at the location of the external force  relative to the suit, but the idea in general is interesting.  The idea of using the measured forces to generate desired trajectories around which a stabilizing controller is designed also appears to be a relatively good idea.  Note that, like any of the model-based control approaches outlined in the exoskeleton literature, or robotics literature for that matter, the feedback linearizing controller in this work is dependent on knowing the model of the robot hardware.

In terms of the control of the overall suit, it was difficult to find any control papers about the Body Extender that covered anything other than upper-body single-arm control of a manipulated object.  The strategy of decoupling the individual limbs from each other, seems somewhat feasible, but is not the best way to approach coordinated limb control.  It was difficult to find literature documenting the specifics of lower-limb control for the Body Extender.  Presumably a similar strategy to that discussed here, with the addition of gravity compensation for the suit, is used.

>>>>

The Body Extender suit is one of the only fully-actuated upper and lower extremity exoskeletons available today.  Clearly, for the task of handling extremely heavy objects, a fully actuated suit may be the best option, but, generally speaking, having the addition weight due to actuating every degree of freedom in the suit is not ideal when considering dynamic operation.

The controllers highlighted in \cite{body_control_2012} do present some novel components.  For example, using a regressor to estimate unknown external forces may be an idea which is directly applicable to a variety of different projects.  Note that in the case of the Body Extender, this regression is dependent on knowing the location of the external force relative to the suit, the manipulator end effectors.  Thus, an extension of this idea, possibly using a lager distribution of sensors to measure external forces, is necessary to extend this approach more generally.

The way that the Body Extender control uses operator-robot measured interaction forces to generate desired trajectories around which a stabilizing controller is designed is generally an idea that we support.  Note that, like any of the model-based control approaches outlined in the exoskeleton literature the feedback linearizing controller in this work is dependent on knowing the model of the robot hardware with a relatively high-level of fidelity.

In terms of the controlling of the overall suit, i.e., locomotion in addition to manipulation, it was difficult to find any papers that discussed more than upper-body single-arm control of a manipulated object for the Body Extender hardware.  The stated strategy of decoupling the individual limbs from each other seems somewhat feasible, but is limited relative to coordinated limb control.

### 2015-09-30 exos/exosText_dualMode.tex
more edits
revision of abstract, summary, and refs
updated ass. and rec. on travers sections

% \begin{document}
\section{Hydraulic Lower Extremity Exoskeleton}

@@

A picture of the Hydraulic Lower Extremity Exoskeleton is presented in Figure \ref{fig:exoSuit}.
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/exoSuit}
  \caption{}
  \vspace{-0.2in}
 \label{fig:exoSuit}   
 \end{figure}

@@

The hydraulic system for each joint is shown in Figure \ref{fig:hydraulicSys}.
A picture of the Hydraulic Lower Extremity Exoskeleton is presented in Figure \ref{fig:exoSuit}.
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/hydraulicSys}
  \caption{}
  \vspace{-0.2in}
 \label{fig:hydraulicSys}   
 \end{figure}

@@

 \begin{figure}[thpb]
\centering
\includegraphics[width=5.in]{exos/figs/hydLowerExrem/dualModeDia}
  \caption{}
  \vspace{-0.2in}
 \label{fig:dualModeDia}   
 \end{figure}
The control design principle for the Hydraulic Lower Extremity Exoskeleton is based on the concept that, for walking gaits, the {\bf stance and swing controllers should be considered separately}.  To this end, the control designers developed an active-passive control method called {\bf Dual-Mode Control}.  Active control is used during stance phases whereas the passive control is used to quickly and freely control the system's legs during swing phases.  Additionally, to handle the inherently discontinuous jumps in the command signals that occur in this dual mode abstraction, a smoothing method is used in conjunction with a pre-transition method that solves for the swing delay due to internal cylinder pressure.  A diagram which highlights the different modes in this control strategy is in Figure \ref{fig:dualModeDia}.

Various components of the dual-mode control strategy, at the hardware level, are presented in Figure \ref{fig:suitDia}.  
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/suitDia}
  \caption{}
  \vspace{-0.2in}
 \label{fig:suitDia}   
 \end{figure}
This figure contains a {\bf gait phase estimation block}, which uses the foot sensors to estimate the stance mode of the system, a {\bf human intention block}, which uses angular encoders and joint force sensors to estimate the motion of the operator, a {\bf torque controller}, and a {\bf gravity compensation} block.  Note that these various components of the control system are used to support the general assumption of the dual mode controller, i.e., 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/torAssist}
  \caption{}
  \vspace{-0.2in}
% \label{fig:suitDia}   
 \end{figure}
 %
 \noindent
where, $GP$ is the gait phase, $GRF$ is the ground reaction force, and $\delta$ is a threshold value.  This equation explicitly states that an active torque will be applied for each leg during stance and that the legs are passive in their respective swing phases.

@@

Using the force sensors on the hydraulic actuators, the {\bf human-robot interaction torques} can be approximated as
\begin{equation}
\hat{\vT}_{hm} = \mM_n(\vq)\ddot{\vq} + \mC_n(\vq,\dot{\vq})\dot{\vq} + \mG_n(\vq) - \hat{\vT}_a, \notag
\end{equation}
where $\mM_n$, $\mC_n$, and $\mG_n$ are the nominal models of the true inertia, centripetal, and gravitational parameters and $\hat{\vT}_a$ represents the measured actuator torques.  The $\mM_n(\vq)\ddot{\vq} + \mC_n(\vq,\dot{\vq})\dot{\vq} + \mG_n(\vq) $ portion of $\hat{\vT}_{hr}$ represents the inverse dynamics component of the closed-loop control system.  The overall joint torque controller, which combines compensation for operator intent as well as balancing of the suit under a gravitational load, is then written \[ {\bf\tau}_\text{active} = K(s) \hat{\vT}_{hm} + \mG_n(\vq),\] where $K(s) = K_p +s K_d,$ i.e., there is closed-loop PD control law operating on the error between the predicted robot dynamics and the measured joint torques.  The block diagram for the active mode controller is shown in Figure \ref{fig:blockDia}. 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/blockDia}
  \caption{}
  \vspace{-0.2in}
 \label{fig:blockDia}   
 \end{figure}
 Note that $\mG_a$ represents the dynamics of the the system and $\mG_a'$ the inverse dynamics model in Figure \ref{fig:blockDia}.

@@

The Hydraulic Lower Extremity Exoskeleton uses a novel scheme to manage transitions between the active and passive control modes.  In particular, the controller applies an explicit smoothing method to address discontinuities in the command torque at phase transitions.  The implementation proposed uses an exponential weighting function of the form
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/weightingForm}
  \caption{}
%  \vspace{-0.2in}
% \label{fig:suitDia}   
 \end{figure}

@@

 \subsection{Opinion}

The design of the Hydraulic Lower Extremity Exoskeleton has a lot going for it.  The best part of this hardware as well as control design is the fact that the creators of it are doing something relatively straightforward.  The actuators are composed of what appears to be off-the-shelf components, and the addition of bypass valves to take advantage of passive dynamics during walking to produce more energy conscious behaviors is a good idea.

One potential drawback of this hardware is the power system.  The documentation in \cite{} was sparse with respect to the exact specifications of the motor and energy storage device which powers the hydraulic system.

The control system for this hardware followed a relatively straight-forward design, effectively creating a ``get out of the way design" using the torque sensed at the joints only.  In the short-term, this type of policy seems to be a good strategy.  The other highlight of the outlined control policy was the smoothing function which was used during phase transitions in the walking gait cycle.  A similar type of smoothing will most likely be beneficial in terms of operator experience for a varity of different operational modes in which different phase transitions occur, e.g., running, jumping, etc.

>>>>

% \begin{document}
\section{Hydraulic Lower Extremity Exoskeleton}
\label{dualMode}
\begin{refsection}[exos/dualMod.bib]

@@

A picture of the Hydraulic Lower Extremity Exoskeleton is presented in Figure \ref{fig:exoSuit}.
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/exoSuit}
  \caption{Exoskeleton robot system.}
%  \vspace{-0.2in}
 \label{fig:exoSuit}   
 \end{figure}

@@

The hydraulic system for each joint is shown in Figure \ref{fig:hydraulicSys}.
A picture of the Hydraulic Lower Extremity Exoskeleton is presented in Figure \ref{fig:exoSuit}.
\begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/hydraulicSys}
  \caption{Dual-mode control}
%  \vspace{-0.2in}
 \label{fig:hydraulicSys}   
 \end{figure}

@@

 \begin{figure}[thpb]
\centering
\includegraphics[width=5.in]{exos/figs/hydLowerExrem/dualModeDia}
  \caption{Locomotion control strategy during gait : Smoothing method ! Active mode ! Pre-transition method ! Passive mode}
%  \vspace{-0.2in}
 \label{fig:dualModeDia}   
 \end{figure}
The control design principle for the Hydraulic Lower Extremity Exoskeleton is based on the concept that, for walking gaits, the {\bf stance and swing controllers should be considered separately}.  To this end, the control designers developed an active-passive control method called {\bf Dual-Mode Control}.  Active control is used during stance phases whereas the passive control is used to quickly and freely control the system's legs during swing phases.  Additionally, to handle the inherently discontinuous jumps in the command signals that occur in this dual mode abstraction, a smoothing method is used in conjunction with a pre-transition method that solves for the swing delay due to internal cylinder pressure.  A diagram which highlights the different modes in this control strategy is in Figure \ref{fig:dualModeDia}.

Various components of the dual-mode control strategy, at the hardware level, are presented in Figure \ref{fig:suitDia}.  
  \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/suitDia}
  \caption{Overall control block diagram.}
%  \vspace{-0.2in}
 \label{fig:suitDia}   
 \end{figure}
This figure contains a {\bf gait phase estimation block}, which uses the foot sensors to estimate the stance mode of the system, a {\bf human intention block}, which uses angular encoders and joint force sensors to estimate the motion of the operator, a {\bf torque controller}, and a {\bf gravity compensation} block.  Note that these various components of the control system are used to support the general assumption of the dual mode controller, i.e., 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/torAssist}
%  \caption{Simplified virtual joint torque control block diagram}
%  \vspace{-0.2in}
% \label{fig:suitDia}   
 \end{figure}
 %
 \noindent
where, $GP$ is the gait phase, $GRF$ is the ground reaction force, and $\delta$ is a threshold value.  This equation explicitly states that an active torque will be applied for each leg during stance and that the legs are passive in their respective swing phases.

@@

Using the force sensors on the hydraulic actuators, the {\bf human-robot interaction torques} can be approximated as
\begin{equation}
\hat{\vT}_{hm} = \mM_n(\vq)\ddot{\vq} + \mC_n(\vq,\dot{\vq})\dot{\vq} + \mG_n(\vq) - \hat{\vT}_a, \notag
\end{equation}
where $\mM_n$, $\mC_n$, and $\mG_n$ are the nominal models of the true inertia, centripetal, and gravitational parameters and $\hat{\vT}_a$ represents the measured actuator torques.  The $\mM_n(\vq)\ddot{\vq} + \mC_n(\vq,\dot{\vq})\dot{\vq} + \mG_n(\vq) $ portion of $\hat{\vT}_{hr}$ represents the inverse dynamics component of the closed-loop control system.  The overall joint torque controller, which combines compensation for operator intent as well as balancing of the suit under a gravitational load, is then written \[ {\bf\tau}_\text{active} = K(s) \hat{\vT}_{hm} + \mG_n(\vq),\] where $K(s) = K_p +s K_d,$ i.e., there is closed-loop PD control law operating on the error between the predicted robot dynamics and the measured joint torques.  The block diagram for the active mode controller is shown in Figure \ref{fig:blockDia}. 
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/blockDia}
  \caption{Simplified virtual joint torque control block diagram.}
%  \vspace{-0.2in}
 \label{fig:blockDia}   
 \end{figure}
 Note that $\mG_a$ represents the dynamics of the the system and $\mG_a'$ the inverse dynamics model in Figure \ref{fig:blockDia}.

@@

The Hydraulic Lower Extremity Exoskeleton uses a novel scheme to manage transitions between the active and passive control modes.  In particular, the controller applies an explicit smoothing method to address discontinuities in the command torque at phase transitions.  The implementation proposed uses an exponential weighting function of the form
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/hydLowerExrem/weightingForm}
%  \caption{Simplified virtual joint torque control block diagram.}
%  \vspace{-0.2in}
% \label{fig:suitDia}   
 \end{figure}

@@

 \subsection{Assessment and Recommendations}

The mechanical as well as control design of the Hydraulic Lower Extremity Exoskeleton was straight-forward and contained many aspects which we support.  In particualr, we supported the use of off-the-shelf components in the hardware design and the addition of bypass valves to take advantage of passive dynamics during walking to produce more energy conscious behaviors.

The control system for this hardware effectively created a ``get out of the way design" using the torque sensed at the joints only.  In the short-term, this type of policy seems to be a good strategy.  The other highlight of the outlined control policy was the smoothing function which was used during phase transitions in the walking gait cycle.  A similar type of smoothing will most likely be beneficial in terms of operator experience for a variety of different operational modes in which different phase transitions occur, e.g., running, jumping, etc.

One potential drawback of this hardware is the power system.  The documentation in \cite{dual_mode_2014} was sparse with respect to the exact specifications of the motor and energy storage device which powers the hydraulic system.  Additionally, in the long-term control design strategy, pure get out of the way control using measurements of the suit only has drawbacks, as mentioned previously.  This strategy will ultimately lead to a ``sensitivity amplification" -like behavior.  This strategy will thus be positively sensitive to user intended motion and negatively sensitive to external disturbances.

The figures as well as hardware and control specifications in this section are taken from \cite{dual_mode_2015}.

\printbibliography[heading=subbibliography]

\end{refsection}

### 2015-09-30 exos/exosText_ihmc.tex
updated ass. and rec. on travers sections

 The design and use of series-elastic actuators is, in general, a good idea for many applications, but may be a problem for the highly dynamic maneuvers required by operators in the TALOS project.  In a large sense, bandwidth is everything in these suits due to the fact that behavior prediction is very complicated.  Simply put, 10Hz control bandwidth at the actuator may be too slow to safely react in dynamically changing environments.

 The other issue with the IHMC is that it only addresses low-level control.  The designers have not specified how to generalize the process of generating torque / joint reference trajectories.  That said, other researchers do explicitly provide methods for generating desired trajectories in what amounts to a mid-level of a hierarchical control scheme.  Thus combining other approaches with the low-level, highly accuracy torque control provided by the novel SEAs presented in this section could offer a potentially good solution at the actuator level.

>>>>

The design of the IHMC MAE is in many ways centered around being able to perform high-fidelity force control.  In general we support the ability to perform force control using a direct measurement (or near direct) of the force/torque at the output of the transmission system.  Series-elastic actuators may provide a solution to this problem for a number of operational scenarios, but caution should be taken when considering this solution for highly dynamic behaviors, as the control bandwidth at the actuator may be too slow to safely react in dynamically changing environments.

One potential other issue with the IHMC control design is that they only appear to address low-level control.  The designers have not specified how to generalize the process of generating torque / joint reference trajectories, in other words there is a lack of information with respect to the typical ``mid-level" of a hierarchical control scheme.  Thus combining other approaches with the low-level, highly accuracy torque control provided by the novel SEAs presented in this section could offer a potentially beneficial control solution.

### 2015-09-30 exos/exosText_roboKnee.tex
updated ass. and rec. on travers sections

 The RoboKnee device is composed of an off-the-shelf knee brace modified with structural pieces to extend the brace and provide actuator attachment points (see Figure~\ref{fig:roboKnee}).
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/roboKnee/roboBrace}
  \caption{RoboKnee design.}
%  \vspace{-0.2in}
 \label{fig:roboKnee}   
 \end{figure}
 The exact length between actuator attachment points is not listed in the literature, but it appears as though the attachment points are at a minimum of one actuator stroke displacement away from each other. Thus, we assume a {\bf minimum moment arm of 30.5 cm} is a reasonable approximation.

@@

 The RoboKnee is a relatively simple system that highlights several intelligent hardware and control design concepts.  The use of a linear ball screw with a series elastic element that enables high-fidelity force control is a very good idea. The idea of using the measured ground reaction forces to produce the desired knee joint torque is also interesting in the sense that it is very simple and relies on a minimal amount of information.

 The overall design of the RoboKnee mechanism is most likely not optimal, as it significantly increases the effective volume of the operator-system leg, but as a prototype, this concept offers interesting benefits.

>>>>

 The RoboKnee device is composed of an off-the-shelf knee brace modified with structural pieces to extend the brace and provide actuator attachment points (see Figure~\ref{fig:roboKnee}).
 \begin{figure}[thpb]
\centering
\includegraphics[width=3.in]{exos/figs/roboKnee/roboBrace}
  \caption{RoboKnee design.}
%  \vspace{-0.2in}
 \label{fig:roboKnee}   
 \end{figure}
 The exact length between actuator attachment points is not listed in the literature, but it appears that the attachment points are at minimum the
stroke displacement (30.5cm) of the actuator away from each other.

@@

 The RoboKnee is single-joint system that highlights several hardware and control design concepts that we generally support.  We again support the use of a linear ball screw with a series elastic element to enable high mechanical efficiency and high-fidelity force control in low to medium speed maneuvers. The concept of using measured ground reaction forces to directly produce desired knee joint torque is interesting in the sense that it is simple and relies on minimal information.

The largest potential issues with the RoboKnee are that the forces of the mechanism and any load carried by the operator are fully transferred through the users musculoskeletal structure.  Additionally, the large moment arm of the actuator significantly increases the effective volume of the operator-system leg, and as such may reduce overall mobility in densely-packed areas.
